---
output: github_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

# AIDA: Analysis of Interval DAta

<!-- badges: start -->
[![R-CMD-check](https://github.com/catarinaploureiro/AIDA/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/catarinaploureiro/AIDA/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/AIDA)](https://CRAN.R-project.org/package=AIDA)
[![CRAN checks](https://badges.cranchecks.info/worst/AIDA.svg)](https://cran.r-project.org/web/checks/check_results_AIDA.html)
[![Github Version](https://img.shields.io/badge/GitHub-v0.1.5-blue)]()
[![License: MIT](https://img.shields.io/badge/license-MIT-blue.svg)](LICENSE)
[![arXiv](https://img.shields.io/badge/arXiv-2604.26769-b31b1b.svg)](https://arxiv.org/abs/2604.26769)
<!-- badges: end -->

## Overview

**AIDA** provides tools for the analysis of interval-valued data, including construction, visualization, robust estimation, and outlier detection. The R package is built around the `intData` class and is designed to support methodological research and applied workflows involving interval-valued data.

AIDA currently includes functionality for:

- Construction of interval-valued datasets
- Interval-valued covariance estimation based on Mallows distance
- Interval Minimum Covariance Determinant (IMCD) estimator
- Robust squared Interval-Mahalanobis distance
- Outlier detection based on robust distances
- Visualization tools for interval data

## Installation

You can install the development version from GitHub:

```{r eval=FALSE}
# install.packages("remotes")
remotes::install_github("catarinaploureiro/AIDA")
```

## Minimal Example

```{r}
library(AIDA)

# Create an intData object from the iris dataset, using the Species column as 
# grouping variable. We also specify the latent distribution as "General" to 
# estimate the parameters based on the microdata.
data(iris)
iris_int <- micro2intData(iris[,1:4], iris$Species, LatentCase = "General")

# Check the parameters of the latent distribution
iris_int@LatentParam

# Compute the classical covariance and correlation matrices
iris_cov <- int_cov(iris_int)
iris_corr <- cov2cor(iris_cov)

# Pairs plot, the lower triangular shows scatter plots of the four variables, 
# while the upper triangular shows the interval correlation matrix.
SYMB.pairs.panels(iris_int, corr = iris_corr, labels = colnames(iris_int))
```

## Vignettes

For a full introduction about the `intData` class (@oliveira2025), see:

```{r eval=FALSE}
vignette("intData_examples", package = "AIDA")
```

For examples on the IMCD estimator and outlier detection based on the robust squared Interval-Mahalanobis distance (@loureiro2026), see:

```{r eval=FALSE}
vignette("IMCD_examples", package = "AIDA")
```

## References